---
title: "Spatio-temporal Interaction Analysis on New York City Shooting Incidents"
author: Fan Li
subtitle: GEO 511 Spatial Data Science Final Project
---

# Introduction

There are hundreds of shooting incidents in New York city on a yearly basis. Despite the statistics of crime events have been declining since 1990 in New York, in the number of shooting incidents was nearly double in 2020 compared to the year before. Even under the circumstances of COVID-19 pandemic, shooting incidents are still happening by a great number. This project explores the following questions: Are these shooting incidents have spatial and temporal connections? Do they happen in particular regions and time periods? What do the spatiotemporal interaction of these events suggests? And why are they clustered spatially and temporally? What could be done given this spatiotemporal interaction? Some basic hypotheses are: There is spatiotemporal interaction among a part of shooting incidents that happen in New York city every year and they could be explained by socioeconomic factors. The spatiotemporal interaction is tested with a statistical method called Knox statistic test. The correlation of spatiotemporal clusters of shooting incidents and socioeconomic variables are calculated to find causalities. As for results and conclusions, there is spatiotemporal interaction among NYC shooting incidents, and these spatiotemporal clusters have causality relationship with socioeconomic environments. 

Key words: Shooting incident, NYC, spatiotemporal interaction, Knox test, socioeconomic factors

# Materials and methods

[~ 200 words]

Narrative: Clear narrative description of the data sources and methods. Includes data from at least two sources that were integrated / merged in R.

Code: The code associated with the project is well organized and easy to follow. Demonstrates mastery of R graphics and functions.

Data: The underlying data are publicly accessible via the web and downloaded/accessed within the Rmd script. If you want to use your own data, you must make it available on a website (e.g. Figshare) so that others are able to re-run your code.

You can do bullets like this:

* The first most important thing
* The second most important thing
* The third most important thing

You can do numbers like this:

1. The first most important thing
2. The second most important thing
3. The third most important thing

See [http://rmarkdown.rstudio.com/](http://rmarkdown.rstudio.com/) for all the amazing things you can do.


Here's my first code chunk.
```{r}
1+2
```

Load any required packages in a code chunk (you may need to install some packages):

```{r, message=F, warning=F}
library(tidyverse)
library(leaflet)
library(kableExtra)
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```

## Download and clean all required data
```{r}
n=20
data=data.frame(x=runif(n,-180,180),
                y=runif(n,-60,60),
                size = runif(n, 5, 20),
                category = factor(
                  sample(letters[1:5], n, replace = TRUE)
                  ),
                value = rnorm(n))
```

```{r, results='asis'}
data %>% 
  slice(1:10) %>% #show only 1:n rows
  kable(digits=2,align="c")%>% #make table and round to two digits
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive")) #apply other formatting
```

Add any additional processing steps here.

# Results

[~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r, fig.width=6, fig.height=3, fig.cap="Map of completely random data"}
m <- leaflet(data) %>% 
  addTiles() %>% 
  addCircleMarkers(~x, ~y, radius = ~size,color = ~as.factor(category)) %>% 
  addPopups(~x[2], ~y[2], "Random popup")
m  # a map with the default OSM tile layer
```


```{r}
data %>% 
  ggplot(aes(x=x,y=y,col=category))+
  geom_point()
```

# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

All sources are cited in a consistent manner
